---
title: "Life expentency of Countries around the world"
author: "Nailah Rawnaq"
date: "1/12/2022"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = F, message = F)
library(tidyverse)
library(tidytext)
library(gganimate)
library(rnaturalearth)

df <- read_csv("Human_life_Expectancy.csv")
df <- 
    df %>% 
    na_if("Not Available")

df <- 
    df %>% 
    pivot_longer(5:34,names_to = "year", values_to = "life_expentancy") %>% 
    mutate(life_expentancy = as.numeric(life_expentancy),
           year = as.numeric(year))
```


## Life expectancy (1990)
```{r}
df %>%
  filter(Level == "National", year == "1990") %>%
  full_join(
    ne_countries(scale = "medium", returnclass = "sf") %>%
      as_tibble() %>%
      select(Country_Code = gu_a3, geometry)
  ) %>%
  ggplot(aes(fill = life_expentancy)) +
  geom_sf(aes(geometry = geometry), col = "white", size = .1) +
  scale_fill_gradient2(
    "Life\nExpectancy",
    high = "#386641",
    mid = "#a7c957",
    low = "#bc4749",
    midpoint = 70,
    na.value = "gray90",
    limits = c(30,90)
  ) +
  theme_light() +
  labs(title = "Plot for life expentancy in 1990",
       x = "Latitude",
       y = "Longditude") +
  theme(
    plot.title = element_text(hjust = .5),
    legend.title = element_text(hjust = .5),
    legend.position = c(.7, .1),
    legend.direction = "horizontal"
  ) 
```

We can see that the life expectancy in the Africa, Indian subcontinent and South Amirica is considerably low than the overall world.


## Life expectancy (2019)
```{r}
df %>%
  filter(Level == "National", year == "2019") %>%
  full_join(
    ne_countries(scale = "medium", returnclass = "sf") %>%
      as_tibble() %>%
      select(Country_Code = gu_a3, geometry)
  ) %>%
  ggplot(aes(fill = life_expentancy)) +
  geom_sf(aes(geometry = geometry), col = "white", size = .1) +
  scale_fill_gradient2(
    "Life\nExpectancy",
    high = "#386641",
    mid = "#a7c957",
    low = "#bc4749",
    midpoint = 70,
    na.value = "gray90", 
    limits = c(30,90)
  ) +
  theme_light() +
  labs(title = "Plot for life expentancy in 2019",
       x = "Latitude",
       y = "Longditude") +
  theme(
    plot.title = element_text(hjust = .5),
    legend.title = element_text(hjust = .5),
    legend.position = c(.7, .1),
    legend.direction = "horizontal"
  ) 
```

We can see that though the Indian sub-continent and Latin america overcome the lower life expediency problem, still African continent is struggling to overcome it.

## Highest life expentancy each year
```{r}
df %>% 
  filter(Level == "National") %>% 
  select(Country, year, life_expentancy) %>% 
  group_by(year) %>% 
  arrange(year, -life_expentancy) %>% 
  mutate(id = row_number()) %>% 
  slice_min(id,n = 15) %>%
  ggplot() +
  geom_segment(aes(x = 14, y = id, 
                   xend = life_expentancy, yend = id, col = Country),size = 5, 
               show.legend = F) +
  geom_text(aes(x = 5, y = id,label = Country),
             hjust = 1,nudge_x = 8, show.legend = F) +
  geom_text(aes(x = life_expentancy, y =  id ,label = life_expentancy),nudge_x = 3) +
  lims(x = c(0,90),
       y = c(15,1)) +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        axis.text.y = element_blank()) +
  transition_states(year) +
  labs(title = "Life expentancy with respect to country.",
       subtitle = "Year: {closest_state}") 
```

This graph show that the Japan remains consistently in top posision with respect to the life expectancy.


## Life expentancy of the country that changed more rapidly
```{r}
temp_df <-
  df %>% 
  filter(Level == "National") %>% 
  nest(data = -Country) %>% 
  mutate(
    lm = map(data, ~lm(data = .x,formula = life_expentancy ~ year)),
    lm = map(lm, tidy),
    lm = map_dbl(lm, ~.x$estimate[2])
  ) %>% 
  slice_max(lm, n = 10)

anim <- 
  temp_df %>% 
  select(-lm) %>% 
  unnest() %>% 
  ggplot(aes(year,life_expentancy, col = Country)) +
  geom_point(show.legend = F) +
  geom_line(show.legend = F) +
  geom_label(aes(label = paste(Country,life_expentancy,sep = ":")),
            nudge_x = 1.8, show.legend = F) +
  geom_text(aes(label = as.factor(year),x = 2020, y = 30), show.legend = F) +
  lims(x = c(NA,2022)) +
  labs(y = "Life Expectancy",
       x = "") +
  transition_reveal(year) 

animation::ani.options(ani.res = 96)
animate(anim,fps = 15, duration = 15)
```

This graph show that there something happen in country Rwanda around 1993-94. On that period the life expectancy is reducing whether for most of the countries it is increasing. We will plot the the country in the world map and find the location of Rwanda as well as Sierra Leone, Uganda, Zambia and Malawi because there is a tendency of falling life expectancy at that period.

```{r}
df %>%
  filter(Level == "National", year == "2019") %>%
  full_join(
    ne_countries(scale = "medium", returnclass = "sf") %>%
      as_tibble() %>%
      select(Country_Code = gu_a3, geometry)
  ) %>%
  mutate(fill =
           ifelse(
             Country %in% c("Rwanda", "Sierra Leone", "Uganda", "Zambia", "Malawi"),
             "yes",
             "no"
           ),
         label = ifelse(fill == "yes",Country,NA)) %>%
  ggplot() +
  geom_sf(
    aes(geometry = geometry, fill = fill),
    col = "white",
    size = .1,
    show.legend = F
  ) +
  labs(title = "World Map marking the countries with anomaly") +
  scale_fill_manual(values = c("#007500", "lightgreen")) +
  coord_sf(xlim = c(-40, 80), ylim = c(-45, 45)) +
  theme(plot.title = element_text(hjust = .5),
        panel.grid = element_blank(),
        panel.background = element_rect(fill = "gray98"))
```

Here we can see that the countries that showed the anomaly are adjacent. So there append something on that period surrounding that place.



















